---
title: "Explore Data"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Claire's root directory
knitr::opts_chunk$set(root.dir="/home/claire/Git/PhenoRepo/")
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Team rotation #1

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(corrplot)
library(ggfortify)
library(reshape2)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(ggfortify)
library(pals)
library(lubridate)
```


-[Essential paper to read?](https://www.nature.com/articles/sdata201828)   
-[Very useful paper about phenocams and chromatic coordinates](https://www.sciencedirect.com/science/article/pii/S2530064417300019)  
-[raw image data if interested](https://data.neonscience.org/data-products/DP1.00033.001)   
-[metadata url](https://daac.ornl.gov/VEGETATION/guides/PhenoCam_V2.html)   
-[GRSM PhenoCam URL](https://phenocam.sr.unh.edu/webcam/sites/NEON.D07.GRSM.DP1.00033/)    
-[GRSM PhenoCam ROI URL with ROI mask file](https://phenocam.sr.unh.edu/webcam/roi/NEON.D07.GRSM.DP1.00033/DB_1000/)   
-[GRSM gcc data csv URL](https://phenocam.sr.unh.edu/data/archive/NEON.D07.GRSM.DP1.00033/ROI/NEON.D07.GRSM.DP1.00033_DB_1000_1day.csv)  
-[Site to download imagedata](https://phenocam.sr.unh.edu/webcam/network/download/)   


```{r data, message=FALSE}
rm(list=ls())

# Claire's Rserver WD
# setwd("/home/claire/Git/PhenoRepo/")

site <- "GRSM"
sites <- c("HARV", "BART", "SCBI", "STEI", "UKFS", "GRSM", "DELA", "CLBJ")

# Individual data--data for every image taken
ind_data <- read_csv(paste0("data/pheno/",site,"/",site,"_individual.csv"))
ind_data <- ind_data[,colSums(is.na(ind_data))<nrow(ind_data)]
# head(ind_data)
# summary(ind_data)

# Daily data--data for images aggregated by day
daily_data <- NULL
for(site in sites){
  
  tmp <- read_csv(paste0("data/pheno/",site,"/",site,"_daily.csv")) %>% 
    filter(image_count>0)
  
  daily_data <- rbind(daily_data,tmp)
}




daily_data <- daily_data[,colSums(is.na(daily_data))<nrow(daily_data)]
# head(daily_data)
# summary(daily_data)

# GDD data -- response variable
gcc_data <- read_csv(paste0("data/pheno/",site,"/",site,"_gccTargets.csv"))
# head(gcc_data)
# summary(gcc_data)

gcol <- 'springgreen4'
rcol <- 'sienna3'
bcol <- 'dodgerblue3'
#doy_pal <- colorRampPalette(kovesi.cyclic_mrybm_35_75_c68(366))(366)
```

**Individual data -- looking at time series plots of data for individual photos**

Histograms of rcc and gcc values
```{r ind_hists}
# names(ind_data)

hist(ind_data$gcc,breaks=50); abline(v=mean(ind_data$gcc,na.rm=T),col='green')
hist(ind_data$rcc,breaks=50); abline(v=mean(ind_data$gcc,na.rm=T),col='red')

```

GCC hist: Sort of multi-modal -- one small peak around 0.10, very high peak around 0.35, and moderate but broader peak around 0.42. Lets try to match these peaks up with times of the year?

RCC: One non-normal peak around 0.37, slight left skew to data.

*Adding horizontal line between plots/plot notes sets*
---

Correlation between the three color channels are in columns in the individual photo data set 
From meta data URL "correlation coefficient (across pixels) between red channel DN and green channel DN, over the ROI"
```{r ind_corr_hists}

# histograms of correlation columns with mean as vertical red line
rgcorrmean <- mean(ind_data$r_g_correl,na.rm=T)
hist(ind_data$r_g_correl,main = paste("mean R-G corr = ", round(rgcorrmean,4))); abline(v=rgcorrmean,col='red')

gbcorrmean <- mean(ind_data$g_b_correl,na.rm=T)
hist(ind_data$g_b_correl,main = paste("mean G-B corr = ", round(gbcorrmean,4))); abline(v=gbcorrmean,col='red')

brcorrmean <- mean(ind_data$b_r_correl,na.rm=T)
hist(ind_data$b_r_correl,main = paste("mean B-R corr = ", round(brcorrmean,4))); abline(v=brcorrmean,col='red')

```

\n
Of correlations between three color channels, highest is R-G (red green), then G-B, then B-R.   
- Are there useful biological takeaways from this? What does it mean to say the G-B and are correlated?
- This also doesn't tell us pos. or neg. correlations. My intuition says that R is negative corr. with B and G, and G and B are positively correlated. 

---

Plot of mean R, G, and B values against time. 
```{r ind_data}
#names(ind_data)

ind_data_melt <- melt(ind_data,id.vars = 'date',measure.vars = c('g_mean',"r_mean","b_mean"))

ggplot(ind_data_melt,aes(x=date,y=value))+
  geom_point(aes(color=variable),alpha=0.05,size=1)+
  facet_wrap(~variable)+
  labs(x="Year", y="Mean channel pixel value")+
  scale_color_manual(values=c(gcol,rcol,bcol))+
  theme_classic()
```

- Pretty clear annual oscillations

---


```{r ind_data2, message=FALSE}
# names(ind_data)
ind_data_melt2 <- melt(ind_data,id.vars = 'solar_elev',measure.vars = c('gcc',"rcc"))

ggplot(ind_data_melt2,aes(x=solar_elev,y=value))+
  geom_point(aes(color=variable),alpha=0.01)+
  geom_smooth(method='lm')+
  facet_wrap(~variable)+
  scale_color_manual(values=c(gcol,rcol))+
  theme_bw()
```
   
Sort of see 3 interesting groups in the green again, maybe related to the three peaks in the histogram

---

**Daily data -- Looking at time series plots of data for photos aggregated by day**.  
This is a plot of r mean, g mean, and b mean across time     
 -'Mean' explanation: "the mean value (for all images passing the selection criteria) of the mean (by image) red channel DN over the ROI"   
  - DN = digital number   
  - "Thus, each pixel in the image is associated with a digital number (“DN”) triplet, with each element in the triplet corresponding to the intensity of one of the colour layers. Therefore, the second step in the image analysis was to read in the images, and associated mask sequence, and to characterize the frequency distribution of the RGB DN triplets within the ROI. We did this separately for each ROI at each site, to produce the “all-image” data files contained in Data Record 3 (see below)."

```{r explore_ind1}
# names(ind_data)

ggplot(daily_data,aes(x=date)) +
  geom_point(aes(y=r_mean),color=rcol,alpha=0.5)+
  geom_point(aes(y=g_mean),color=gcol,alpha=0.5)+
  geom_point(aes(y=b_mean),color=bcol,alpha=0.5)+
  geom_vline(xintercept = as.Date('2017-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2018-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2019-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2020-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2021-01-01'),color='grey50',alpha=0.5)+
  labs(y="mean daily pixel value for each channel")+
  theme_classic()
```

- Looks like red peaks around November -- severe drop off right as green peaks.  
- Blue values generally much lower

---

**Plot gcc and rcc over time**.   
cc = chromatic coordinate = " the ratio of the amount of one primary color to the total amount of all three necessary to reproduce a given color."
```{r explore_ind2}
ggplot(daily_data,aes(x=date))+
  geom_vline(xintercept = as.Date('2017-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2018-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2019-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2020-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2021-01-01'),color='grey50',alpha=0.5)+
  geom_point(aes(y=rcc_90),color=rcol,alpha=0.5)+
  geom_point(aes(y=gcc_90),color=gcol,alpha=0.5)+
  theme_classic()
```

---

**Look more at gcc through time**
```{r more_gcc}

gcc_df <- daily_data %>% 
  select(date,year,doy,
         midday_gcc,
         g_mean,
         gcc_mean,gcc_std,
         gcc_50,gcc_75,gcc_90)

ggplot(gcc_df,aes(x=date))+
  geom_point(aes(y=gcc_90,color=as.factor(year)))+
  theme_classic()
```

---

**Maximum daily solar elevation vs gcc_90**
```{r explore_ind3}

# Look at max solar elev through time
mse_plot <- ggplot(daily_data,aes(x=date,y=max_solar_elev))+
  geom_vline(xintercept = as.Date('2017-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2018-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2019-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2020-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2021-01-01'),color='grey50',alpha=0.5)+
  geom_smooth()+
  # geom_line()+ # Weird where missing values
  theme_classic()

gcc_plot <- ggplot(daily_data,aes(x=date,y=gcc_90))+
  geom_vline(xintercept = as.Date('2017-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2018-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2019-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2020-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2021-01-01'),color='grey50',alpha=0.5)+
  geom_line(color=gcol)+
  theme_classic()

plot_grid(mse_plot,gcc_plot,ncol = 1)

# Try plotting together
ggplot(daily_data,aes(x=date))+
  geom_vline(xintercept = as.Date('2017-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2018-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2019-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2020-01-01'),color='grey50',alpha=0.5)+
  geom_vline(xintercept = as.Date('2021-01-01'),color='grey50',alpha=0.5)+
  geom_point(aes(y=log(max_solar_elev)/10),size=0.5,color='goldenrod3')+
  geom_line(aes(y=gcc_90),color=gcol)+
  labs(y="",title="GCC_90 and log(solar elev)/10 vs. time",subtitle = "Max solar elev transformed to compare peaks more easily")+
  # geom_line()+ # Weird where missing values
  theme_classic()

ggplot(daily_data,aes(y=gcc_90,x=max_solar_elev))+
  geom_point(aes(color=doy),alpha=0.5)+
  scale_color_gradientn(colors=doy_pal)+
  geom_smooth(method='lm')+
  theme_classic()+
  labs(color="Day of year")

summary(lm(gcc_90~max_solar_elev+doy,data = daily_data))

```

---

Correlation plots
```{r corrplots}
names(daily_data)
corr_df <- daily_data %>% 
  select(1,2,6:27)
summary(corr_df)

corrplot(cor(corr_df[,2:ncol(corr_df)]))
```

- Strong negative correlations between green variables and red variables   
- Strong positive correlation with max_solar_elev and green variables   

---

**PCA Plots**
```{r pcaplots}
# names(daily_data)

pca_df <- daily_data %>% 
  dplyr::select(year,doy,
                # midday_r,midday_g,midday_b,
                # r_mean,g_mean,b_mean,
                # gcc_mean,rcc_mean,
                gcc_90,rcc_90,
                max_solar_elev)

pca_res <- prcomp(pca_df, scale. = TRUE)

autoplot(pca_res, data = daily_data, 
         colour = 'year',
         # colour='grey50',
         loadings = TRUE, loadings.label=TRUE,
         loadings.colour='grey50',loadings.label.colour='darkred',
         loadings.label.size = 5,alpha = 0.5)+
  theme_classic()+
  scale_color_viridis(option = "D")

```

# Plot PC1/PC2 against other variables

```{r more_pca}
PCs <- as.data.frame(pca_res$x)
PCs12_df <- cbind(PC1=PCs$PC1,PC2=PCs$PC2, daily_data)
# summary(PCs12_df)

ggplot(PCs12_df,aes(x=PC1,y=gcc_90))+
  geom_point(aes(color=doy),alpha=0.5)+
  scale_color_gradientn(colors=doy_pal)+
  geom_smooth(method="lm",se=T)+
  # geom_smooth(se=FALSE,color='red')+
  theme(legend.position = "none")+
  theme_classic()

```


https://www.sciencedirect.com/science/article/pii/S0378112716302298

```{r}
targets <- read_csv("phenology-targets.csv.gz") %>% 
  mutate(doy = yday(targets$time),
         year=str_sub(time,1,4))

siteinfo <- read_csv("data/pheno/siteinfo.csv")

temp <- read_csv("./data/drivers/neon/GDDandtemp_allsites.csv") %>% 
  select(date,siteID,daily_mean,midday_mean,GDDdaily,GDDdays,GDDlogic,GDDtotal) 
daily_data$siteID <- str_sub(daily_data$midday_filename,10,13)
df <- base::merge(daily_data,temp,by.x=c("date","siteID"),by.y = c("date","siteID"))

# create DF with lagged GDD daily
lag_df <- NULL
lag<-15
for(site in sites){
  
  tmp <- filter(df,siteID==site)
  GDDdaily_lag = c(tmp$GDDdaily[1:(nrow(tmp)-lag)])
  tmp2 <- cbind(tmp[(lag+1):nrow(tmp),],GDDdaily_lag)
  lag_df <- rbind(lag_df,tmp2)
  }

summary(lm(gcc_90~GDDdaily,df))
summary(lm(gcc_90~GDDdaily_lag,lag_df))

ggplot(df,aes(x=GDDtotal,y=gcc_90))+
  geom_point()+
  facet_wrap(~siteID)+
  geom_smooth()+
  theme_classic()

ggplot(df,aes(x=GDDdaily,y=gcc_90))+
  geom_point()+
  facet_wrap(~siteID)+
  geom_smooth(method="lm")+
  theme_classic()

ggplot(df,aes(x=doy,y=gcc_90))+
  geom_point(aes(color=doy))+
  facet_wrap(~siteID)+
  geom_smooth()+
  theme_classic()

site = "GRSM"
ginc <- siteinfo$field_avg_grean_increase_doy[siteinfo$siteID==site]
gdec <- siteinfo$field_avg_green_decrease_doy[siteinfo$siteID==site]
gmax <- siteinfo$field_avg_green_max_doy[siteinfo$siteID==site]
gmin <- siteinfo$field_avg_green_min_doy[siteinfo$siteID==site]
gndays <- siteinfo$field_avg_number_of_green_days[siteinfo$siteID==site]
gcc_min <- min(targets$gcc_90[targets$siteID==site],na.rm=T)-0.01
gcc_max <- max(targets$gcc_90[targets$siteID==site],na.rm=T)+0.01

ggplot(subset(targets,siteID==site), aes(x=doy,y=gcc_90))+
  geom_rect(aes(xmin=ginc,xmax=ginc+gndays,ymin=gcc_min,ymax=gcc_max),fill='grey90',alpha=0.25)+
  geom_point(aes(color=year))+
  geom_smooth()+
  geom_vline(xintercept = ginc,color='green')+
  geom_text(aes(x=ginc, label="green increase DOY", y=gcc_max), angle=90, vjust=0,hjust=1)+ 
  geom_vline(xintercept = gmax,color='forestgreen')+ # This shows that the windy year really skews this value at GRSM
  geom_text(aes(x=gmax, label="green max DOY", y=gcc_max), angle=90, vjust=0,hjust=1)+ 
  geom_vline(xintercept = gdec,color='darkgreen')+ 
  geom_text(aes(x=gdec, label="green decrease DOY", y=gcc_max), angle=90, vjust=0,hjust=1)+ 
  geom_vline(xintercept = gmin,color='indianred4')+
  geom_text(aes(x=gmin, label="green min DOY", y=gcc_max), angle=90, vjust=0,hjust=1)+ 
  scale_y_continuous(expand=c(0,0))+
  ggtitle("Figure to show example DOY values in siteinfo.csv")+
  theme_classic()
ggsave("data/pheno/siteinfo_doyvals_ex.png")
```

```{r}
rm(list=ls())

targets <- read_csv("phenology-targets.csv.gz") 

targets <- targets %>% 
  mutate(doy = yday(targets$time),
         year=str_sub(time,1,4))

siteinfo <- read_csv("data/pheno/siteinfo.csv")
temp <- read_csv("data/drivers/neon/GDDandtemp_allsites.csv")
colnames(temp)[1]<-"time"

df <- merge(temp,targets)

ggplot(subset(targets,siteID=="GRSM"), aes(x=doy,y=gcc_90))+
  geom_vline(xintercept=80,color="grey70",linetype=3,size=0.9)+
  geom_vline(xintercept=172,color="grey70",linetype=3,size=0.9)+
  geom_vline(xintercept=264,color="grey70",linetype=3,size=0.9)+
  geom_vline(xintercept=355,color="grey70",linetype=3,size=0.9)+
  geom_point(size=3,alpha=0.5,color="springgreen4")+
  labs(x="Day of year",y="GCC 90")+
  scale_x_continuous(expand=c(0.01,0))+
  theme_classic(base_size = 15)
ggsave("gccdataexample.png",width=10,height=5,dpi = 400)

colnames(temp)[1]<-"time"
  
ggplot(df, aes(x=gcc_90,y=,midday_mean))+
  geom_point(aes(color=siteID),alpha=0.5)+
  labs(x="GCC 90",y="Daily mean temperature",color="Site ID")+
  geom_smooth(method="lm")+
  scale_y_continuous(limits = c(0,30),expand=c(0,0))+
  theme_classic(base_size = 20)

ggplot(df,aes(x=gcc_90,y=GDDdaily))+
  geom_point(aes(color=siteID),alpha=0.5)+
  labs(x="GCC 90",y="GDD Daily",color="Site ID")+
  geom_smooth(method="lm")+
  facet_wrap(~siteID)
  theme_classic(base_size = 20)

ggplot(df,aes(x=GDDdaily,y=gcc_90))+
  geom_point(aes(color=siteID),alpha=0.5)+
  labs(x="GCC 90",y="GDD Daily",color="Site ID")+
  theme_classic(base_size = 20)

ggplot(df,aes(x=gcc_90,y=GDDtotal))+
  geom_point(aes(color=siteID),alpha=0.5)+
  labs(x="GCC 90",y="GDD Daily",color="Site ID")+
  theme_classic(base_size = 20)

ggplot(df,aes(x=gcc_90,y=GDDdays))+
  geom_point(aes(color=siteID),alpha=0.5)+
  labs(x="GCC 90",y="GDD Daily",color="Site ID")+
  theme_classic(base_size = 20)

trendline(df$daily_mean,df$gcc_90)
trendline(df$GDDdaily,df$gcc_90)


summary(lm(gcc_90~GDDdaily,df))
```






```{r}
pca_df <- df[, unlist(lapply(df, is.numeric))] %>% 
  filter(complete.cases(.)) %>% 
  select(gcc_90,gcc_sd,doy,GDDdaily,midday_mean,daily_mean,year)
  
corrplot(cor(pca_df), method = c("number"), type = 'upper',add=T)

pca_res <- prcomp(pca_df, scale. = TRUE)

autoplot(pca_res, data = pca_df,
         loadings = TRUE, loadings.label=TRUE,
         loadings.colour='grey50',loadings.label.colour='darkred',
         loadings.label.size = 5,alpha = 0.75)+
  theme_classic()+
  scale_color_viridis(option = "D")

PCs <- as.data.frame(pca_res$x)
PC_df <- cbind(PC1=PCs$PC1,PC2=PCs$PC2, df)

trendline(PC_df$PC1,PC_df$log_max_ls)
ggplot(PC_df,aes(x=PC1,y=log_max_ls))+
  geom_point()+
  #scale_color_viridis()+
  geom_smooth(method="lm",se=T)+
  # geom_label(aes(label=species))+
  theme_classic(base_size=20)+
  # geom_text(aes(x=-1.8,y=8,label="y = 0.707x + 2.660"),size=8)+
  # geom_text(aes(x=-2.1,y=7.5,label="R^2=0.704, p<0.0001"),size=5.5)+
  labs(x="PC1 (27.6%)", y = "Log(maximum lifespan)")

```


###### PRESENTATION FIGURE CODE #######
```{r prezFigCode}
ggplot(subset(targets,siteID=="GRSM"), aes(x=doy,y=gcc_90))+
  geom_vline(xintercept=80,color="grey70",linetype=3,size=0.9)+
  geom_vline(xintercept=172,color="grey70",linetype=3,size=0.9)+
  geom_vline(xintercept=264,color="grey70",linetype=3,size=0.9)+
  geom_vline(xintercept=355,color="grey70",linetype=3,size=0.9)+
  geom_point(size=3,alpha=0.5,color="springgreen4")+
  labs(x="Day of year",y="Green chromatic coordinate")+
  scale_x_continuous(expand=c(0.01,0))+
  theme_classic(base_size = 15)
ggsave("gccdataexample.png",width=10,height=5,dpi = 400)

```



```{r data}

ind <- read_csv("data/pheno/GRSM/GRSM_individual.csv") 
# gcc = ind$g_mean[1]/(sum(ind$g_mean[1]+ind$r_mean[1]+ind$b_mean[1]))

targets <- read_csv("data/pheno/GRSM/GRSM_gccTargets.csv") %>% 
  filter(!is.na(gcc_90))

```

### MEAN OF WHAT
So where does gcc_90 come from in the daily df?
```{r find_gcc90}

ind_10 <- ind %>% 
  filter(solar_elev>10)

midday_gcc <- ind %>%  
  group_by(date) %>% 
  summarize(#meangcc = mean(gcc,na.rm=T),
            # SE1_gcc_90 = as.numeric(quantile(gcc[solar_elev>1],probs=0.9,na.rm=T)),
            SE10_gcc_90 = round(as.numeric(quantile(gcc[solar_elev>10],probs=0.9,na.rm=T)),5),
            SE20_gcc_90 = as.numeric(quantile(gcc[solar_elev>20],probs=0.9,na.rm=T)),
            SE30_gcc_90 = as.numeric(quantile(gcc[solar_elev>30],probs=0.9,na.rm=T)))

midday_gcc <- merge(midday_gcc,targets, by.x='date',by.y='time')
colnames(midday_gcc)[6]<-'target_gcc_90'
midday_gcc$equal = midday_gcc$SE10_gcc_90==midday_gcc$target_gcc_90

a <- read_csv("./data/pheno/allsites_gccTargets.csv")
```


