---
title: "Seasonal Arima"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir="/home/claire/Git/PhenoRepo/")
```

Load packages
```{r packages}

rm(list=ls())

# remotes::install_github("eco4cast/neon4cast")

library(tidyverse)
library(fpp)
library(forecast)
library(neon4cast)
library(tsibble)
library(fable)
library(scoringRules)
library(zoo)
library(reshape2)

```

Get data
```{r data}
full_data <- read_csv("./data/pheno/allsites_gccTargets.csv") %>% 
  filter(time <= as.Date('2021-03-01'))

download.file("https://data.ecoforecast.org/targets/phenology/phenology-targets.csv.gz","data/pheno/phenology-targets.csv.gz")
targets <- read_csv("data/pheno/phenology-targets.csv.gz")

sites <- unique(full_data$siteID)
```

Fix issue where data begins with NA, but leave other NAs (mainly @ UKFS)
```{r}

full_data_fixed <- NULL

for(s in 1:length(sites)){
  
  site <- sites[s]
  tmp <- filter(full_data,siteID==site)

  while(is.na(tmp$gcc_90[1])){
    
    tmp <- tmp[2:nrow(tmp),]
    
  }
    
  full_data_fixed <- rbind(full_data_fixed,tmp)

  rm(tmp)
}

full_data <- full_data_fixed

```


# Fill in other NA values (SHADY MOVES RIGHT NOW)
```{r fillNA}

gcc_noNA <- rollapply(full_data$gcc_90, width = 20, function(x){
  
  mean(x,na.rm = T)
  
  }, fill = NA)

full_data$gcc_noNA <- gcc_noNA
full_data$gcc_old <- full_data$gcc_90
full_data$gcc_90 <- ifelse(is.na(full_data$gcc_90),full_data$gcc_noNA,full_data$gcc_90)

full_data$gcc_90[is.nan(full_data$gcc_90)]<-NA

ggplot(full_data,aes(x=time,y=gcc_90))+
  geom_point()+
  facet_wrap(~siteID,scales='free')
```

Seasonal ARIMA by site
```{r sarima}

all_forecast <- NULL

for(s in 1:length(sites)){
  print(sites[s])
  
  # filter data frame for current site
  tmp_df = full_data %>% 
    filter(siteID == sites[s]) %>% 
    arrange(time) %>% 
    select(1:4)
  
  first_date <- as.Date(min(tmp_df$time))
  year <- as.numeric(str_sub(first_date,1,4))
  doy <- as.numeric(strftime(first_date, format = "%j"))
  
  # convert gcc to time series data
  tmp_ts = ts(tmp_df$gcc_90, start=c(year,doy), deltat = 1/365)
  tmp_arima <- auto.arima(tmp_ts,trace=T)
  assign(paste0(sites[s],"_arima"),tmp_arima)
  
  tmp_forecast <- forecast::forecast(tmp_arima,h=35,bootstrap=T)
  assign(paste0(sites[s],"_4cast"),tmp_forecast)

  tmp_4cast_df <- data.frame(time = seq(as.Date(max(tmp_df$time)), by="day",length.out = 35),
                            siteID = sites[s],
                            gcc_90 = as.numeric(round(tmp_forecast$mean,4)),
                            gcc_sd = as.numeric(round(tmp_forecast$mean-tmp_forecast$lower[2],4))) 
  all_forecast <- rbind(all_forecast,tmp_4cast_df)
  
  rm(tmp_df,tmp_4cast_df,tmp_arima,tmp_forecast,year,doy,tmp_ts,first_date)

  }

all_data <- rbind(full_data[,1:4],all_forecast) %>% arrange(siteID,time)
save(list=ls(all.names=T),file="R/ARIMA_20210301.Rdata")
```

```{r}

ggplot(all_data,aes(x=time,y=gcc_90))+
  geom_line()+
  facet_wrap(~siteID)+
  theme_bw()


for(s in 1:length(sites)){
  
  tmp_4cast <- eval(as.name(paste0(sites[s],"_4cast")))
  plot(tmp_4cast,main=sites[s])
  
}
```

7.1.4 Phenology
    time: YYYY-MM-DD
    siteID: NEON code for site
    ensemble or statistic: if ensemble then integer value for forecast replicate within the year and month (i.e. ensemble member or MCMC sample); if statistic then either required to be the string mean or sd (see note below about statistic column).
    forecast: set as 1 for each row (1 = variables were forecasted; a 0 would designate a hindcast which does not apply to submissions to the challenge)
    data_assimilation: set as 0 for each row (0 = no data assimilation occurred because it is a forecast)
    gcc_90: green chromatic coordinate

Scoring
```{r scoring}
example <- readr::read_csv("https://data.ecoforecast.org/forecasts/phenology/phenology-2021-02-23-UCSC_P_EDM.csv")

score_df <- melt(all_forecast,id.vars = c("time","siteID"),
                 variable.name = "statistic",
                 value.name = "gcc_90") %>% 
  mutate(forecast=1,
         obs_flag=2,
         data_assimilation=0) %>% 
  select(time,siteID,obs_flag,forecast,data_assimilation,statistic,gcc_90) %>% 
  as_tibble()

# replace gcc_90 and gcc_sd with mean and sd, resp. 
score_df$statistic <- ifelse(score_df$statistic=="gcc_90","mean","sd")
score_df$statistic <- as.character(score_df$statistic)

# class(score_df) <- class(example) # Didn't work

# This only seems to work when we want to score relative to PhenoCam data that has already been observed
scores <- neon4cast::score(score_df,theme=c("phenology"))

```




